---
title: "Cyclist_Capstone"
author: "Hima"
date: '2022-05-23'
---

```{r setup, include= TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

### Install packages
```{r}
#install.packages("tidyverse")
#install.packages("lubridate")
#install.packages("janitor")
library(tidyverse)
library(lubridate)
library(janitor)
```

### Importing Data (CSV files)

```{r}
Tripdata_Jan21 <- read_csv('Tripdata_Jan21.csv')
Tripdata_Feb21 <- read_csv('Tripdata_Feb21.csv')
Tripdata_Mar21 <- read_csv('Tripdata_Mar21.csv')
Tripdata_Apr21 <- read_csv('Tripdata_Apr21.csv')
Tripdata_May21 <- read_csv('Tripdata_May21.csv')
Tripdata_Jun21 <- read_csv('Tripdata_Jun21.csv')
Tripdata_Jul21 <- read_csv('Tripdata_Jul21.csv')
Tripdata_Aug21 <- read_csv('Tripdata_Aug21.csv')
Tripdata_Sep21 <- read_csv('Tripdata_Sep21.csv')
Tripdata_Oct21 <- read_csv('Tripdata_Oct21.csv')
Tripdata_Nov21 <- read_csv('Tripdata_Nov21.csv')
Tripdata_Dec21 <- read_csv('Tripdata_Dec21.csv')
```


### Checking the COLUMN Names and no. of columns of the files imported

```{r}
colnames(Tripdata_Jan21)
colnames(Tripdata_Feb21)
colnames(Tripdata_Mar21)
colnames(Tripdata_Apr21)
colnames(Tripdata_May21)
colnames(Tripdata_Jun21)
colnames(Tripdata_Jul21)
colnames(Tripdata_Aug21)
colnames(Tripdata_Sep21)
colnames(Tripdata_Oct21)
colnames(Tripdata_Nov21)
colnames(Tripdata_Dec21)
```

### Check the Data Type of each Column of all the files
```{r}
str(Tripdata_Jan21)
str(Tripdata_Feb21)
str(Tripdata_Mar21)
str(Tripdata_Apr21)
str(Tripdata_May21)
str(Tripdata_Jun21)
str(Tripdata_Jul21)
str(Tripdata_Aug21)
str(Tripdata_Sep21)
str(Tripdata_Oct21)
str(Tripdata_Nov21)
str(Tripdata_Dec21)
```

### Checking Column Names of all the files
```{r}
compare_df_cols(Tripdata_Jan21, Tripdata_Feb21, Tripdata_Mar21, Tripdata_Apr21, Tripdata_May21, Tripdata_Jun21, Tripdata_Jul21, Tripdata_Aug21, Tripdata_Sep21, Tripdata_Oct21, Tripdata_Nov21, Tripdata_Dec21, return  = "mismatch")

```

### Merge all files (12 months data) to single file

```{r}
Tripdata_2021 <- rbind(Tripdata_Jan21, Tripdata_Feb21, Tripdata_Mar21, Tripdata_Apr21, Tripdata_May21, Tripdata_Jun21, Tripdata_Jul21, Tripdata_Aug21, Tripdata_Sep21, Tripdata_Oct21, Tripdata_Nov21, Tripdata_Dec21)

```

### Finding missing values in Columns (https://www.codingprof.com/3-ways-to-find-columns-with-nas-in-r-examples/)
```{r}
colSums(is.na(Tripdata_2021))
```

### Remove/Drop missing data from data frame
```{r}
Tripdata_2021 <- Tripdata_2021 %>% drop_na()
dim(Tripdata_2021)
```

### Checking for duplicate data
```{r}
duplicate_check <- distinct(Tripdata_2021, ride_id, .keep_all = TRUE)
dim(duplicate_check)

```

### removing unwanted columns 'start_lat', 'start_lng', 'end_lat', 'end_lng' from data frame
```{r}
Tripdata_2021 <- Tripdata_2021 %>% select(-c('start_lat', 'start_lng', 'end_lat', 'end_lng'))
```

### Exporting the file in CSV format
```{r}
write.csv(Tripdata_2021, "Tripdata_2021.csv")

```


